---
title: "Adverse Event Data"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{aeplots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(printr)
library(dplyr)
library(tidyverse)
library(tidyr)
```

## Introduction
The vignette documents how to use the functions in `aeplots` to create tables and plots for analysing adverse event data in clinical trials.
```{r setup, message=FALSE}
library(aeplots)
```

## Input data
### 2 treatment arms
To generate tables and plots to summarise adverse events in clinical trials, the R package requires the input dataset in data frame format. A sample dataset with two treatment arms is shown below:
```{r}
data(df2)
head(df2, 10)
```

Note that for each participant who did not experience any adverse events, a row should be included in the input dataset stating his/her `id`, `arm`, `date_rand`, `last_visit` and `variables` to be included in the model. `adverse_event`, `body_system_class` and `severity` column should be specified as `NA`. For example:
```{r}
tail(df2 %>% 
       add_row(id=2110, arm="Intervention", ae_pt=NA, aebodsys=NA, severity=NA, date_rand=ymd("2015-05-22"),
               last_visit=ymd("2016-04-22"), variable1=2, variable2=1500), 1)
```

We would need to convert the body_system_class and severity columns to factors.
```{r}
df2$aebodsys <- as.factor(df2$aebodsys)
# for severity to be presented in ascending order, we order the levels
df2$severity <- ordered(df2$severity, c("Mild", "Moderate", "Severe"))
```

Note that for all the functions below, you would need to specify the column names corresponding to each variable needed unless you rename the column names of your dataset to the default column names as specified in the sample dataset of the documentation. Do refer to the detailed documentation of each function by typing:
```{r, results="hide"}
help(aetable)
```


### More than 2 treatment arms
The functions `aetable`, `aeseverity`, `aestacked` and `aebar` can take up to 4 treatment arms. A sample dataset with 3 treatment arms:
```{r}
data(df3)
head(df3, 10)
df3$aebodsys <- as.factor(df3$aebodsys)
df3$severity <- ordered(df2$severity, c("Mild", "Moderate", "Severe"))
```

## `aetable` function
### 2 treatment arms
`aetable` plots a table of AE summary by body system class and arm. It contains the total number of participants at risk per arm ($N_{1}$, $N_{2}$), frequency (N), proportion (%), total number of events (n), incidence rate (IR), number of adverse events per participant (mean & SD) and treatment effect estimate with its 95% confidence interval (CI). To include additional covariates besides arm in the model, specify a vector of variable names in the `variables` argument.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        treatment_effect_estimate=TRUE, variables = c("variable1", "variable2"))
```

We can specify the model used to estimate the treatment effect and 95% CI via the `model` argument. The available model options are `Poisson (rate)`, `Poisson (count)`, `Negative Binomial (rate)`, `Negative Binomial (count)`, `Binomial (logit)`, `Binomial (log)` and `Binomial (identity)`.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        treatment_effect_estimate=TRUE, model="Binomial (logit)", variables = c("variable1", "variable2"))
```

We can specify IR per number of person via the `IR_per_person` argument. The default is IR per 100 person.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        IR_per_person=1000)
```

We can specify the number of decimal places for the proportions, IR, mean and SD columns as well as the number of significant figures for the treatment effect estimate and 95% CI.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        treatment_effect_estimate=TRUE, variables = c("variable1", "variable2"), proportions_dp=2, IR_dp=2,
        mean_dp=2, SD_dp=2, estimate_sf=4, CI_sf=4)
```

We can choose to drop the treatment effect estimate and 95% CI by specifying `treatment_effect_estimate=FALSE`.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        treatment_effect_estimate=FALSE)
```

We can also choose to drop the mean column by specifying `mean=FALSE`.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        treatment_effect_estimate=FALSE, mean=FALSE)
```

### More than 2 treatment arms
`aetable` does not include the treatment effect estimates and their 95% CIs for datasets with more than 2 treatment arms. To change the labels of control and interventions in the table, specify the label for control in the argument `control_name` and specify the intervention labels in the argument `intervention_names`. Note that intervention_names should be specified using the same order as the interventions specified in intervention_levels.
```{r, fig.dim=c(8, 6)}
aetable(df3, body_system_class="aebodsys", control="Placebo", control_name="No Drug",
        intervention_levels=c("Intervention 1", "Intervention 2"), intervention_names=c("Drug A", "Drug B"))
```

## `aeseverity` function
### 2 treatment arms
`aeseverity` plots a table of frequencies and proportions of events by severity categories. We specify the number of decimal places for proportions via the argument `proportions_dp`.
```{r, fig.dim=c(8, 6)}
aeseverity(df2, arm_levels=c("Intervention", "Placebo"), body_system_class="aebodsys", proportions_dp=2)
```

We can choose to drop the number of events column by specifying `n_events=FALSE`.
```{r, fig.dim=c(8, 6)}
aeseverity(df2, arm_levels=c("Intervention", "Placebo"), body_system_class="aebodsys", n_events=FALSE)
```

### More than 2 treatment arms
```{r, fig.dim=c(8, 6)}
aeseverity(df3, arm_levels=c("Intervention 1", "Intervention 2", "Placebo"), 
           arm_names=c("Drug A", "Drug B", "No Drug"), body_system_class="aebodsys", n_events=FALSE)
```

## `aedot` function
`aedot` plots a dot plot proportions alongside treatment effect estimates with accompanying 95% confidence interval to visualise AE and harm profiles in two-arm randomised controlled trials.
```{r, message=FALSE, warning=FALSE, fig.dim=c(8, 6)}
aedot(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention")
```

Similar to `aetable`, we can specify the model used to estimate the treatment effect and 95% CI via the `model` argument. The available model options are `unadjusted (RR)`, `unadjusted (RD)`, `Poisson (rate)`, `Poisson (count)`, `Negative Binomial (rate)`, `Negative Binomial (count)`, `Binomial (logit)`, `Binomial (log)` and `Binomial (identity)`. The default model is `unadjusted (RR)`.
```{r, message=FALSE, warning=FALSE, fig.dim=c(8, 6)}
aedot(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention", model="unadjusted (RD)")
```

We can change the colours of the dots on the percentage of participants plot through the `dot_colours` argument.
```{r, message=FALSE, warning=FALSE, fig.dim=c(8, 6)}
aedot(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention", 
      dot_colours=c("#F4B800", "#97D800"))
```

## `aestacked` function
### 2 treatment arms
`aestacked` plots a stacked bar chart of proportions for each body system class by arm and maximum severity. If severity is not ordered, it is required to specify the levels of severity in ascending order in the argument `severity_levels`.
```{r, fig.dim=c(8, 6)}
aestacked(df2, body_system_class="aebodsys", arm_levels=c("Intervention","Placebo"), 
          severity_levels=c("Mild", "Moderate", "Severe"))
```

We can change the colour for each severity by specifying a vector of colour codes corresponding to each severity level in the `severity_colours` argument. 
```{r, fig.dim=c(8, 6)}
aestacked(df2, body_system_class="aebodsys", arm_levels=c("Intervention","Placebo"), 
          severity_levels=c("Mild", "Moderate", "Severe"), severity_colours=c("#D0EE11", "#48B5C4", "#115F9A"))
```

### More than 2 treatment arms
```{r, fig.dim=c(8, 6)}
aestacked(df3, body_system_class="aebodsys", arm_levels=c("Intervention 1","Intervention 2", "Placebo"),
          arm_names=c("Drug A","Drug B", "No Drug"), severity_levels=c("Mild", "Moderate", "Severe"))
```

## `aebar` function
### 2 treatment arms
`aebar` plots a bar chart for number of events reported per participant.
```{r, fig.dim=c(8, 6)}
aebar(df2, arm_levels=c("Intervention","Placebo"), adverse_event="ae_pt", facets=FALSE)
```

We can change the colour representing each treatment arm by specifying a vector of colour codes in the `arm_colours` argument according to the order of arm levels specified in the arm_levels argument.
```{r, fig.dim=c(8, 6)}
aebar(df2, arm_levels=c("Intervention","Placebo"), adverse_event="ae_pt", arm_colours=c("#FFB55A", "#B2E061"), facets=FALSE)
```

### More than 2 treatment arms
```{r, fig.dim=c(8, 6)}
aebar(df3, arm_levels=c("Intervention 1","Intervention 2", "Placebo"), arm_names=c("Drug A","Drug B", "No Drug"),
      adverse_event="ae_pt", facets=FALSE)
```

We can plot a bar chart for each treatment arm in separate facets by specifying `facets=TRUE`.
```{r, fig.dim=c(8, 6)}
aebar(df3, arm_levels=c("Intervention 1","Intervention 2", "Placebo"), arm_names=c("Drug A","Drug B", "No Drug"),
      adverse_event="ae_pt", facets=TRUE)
```

## `aevolcano` function
`aevolcano` plots a volcano plot of treatment effect estimate against the p-value on the -log10 scale. 
```{r, fig.dim=c(8, 6)}
aevolcano(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention")
```

We can specify the model used to estimate the treatment effect and 95% CI via the `model` argument. The available model options are `unadjusted (RD)`, `unadjusted (RR)`, `unadjusted (OR)`, `unadjusted (IRR)`, `Poisson (rate)`, `Poisson (count)`, `Negative Binomial (rate)`, `Negative Binomial (count)`, `Binomial (logit)`, `Binomial (log)` and `Binomial (identity)`. The default model is `unadjusted (RD)`.
```{r, fig.dim=c(8, 6)}
aevolcano(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention", model="Binomial (logit)",
          variables=c("variable1", "variable2"))
```

Labels are added to events with p-values less than 0.1 on default. We can change the p-value cutoff for which labels are added to events through the `p_value_cutoff` argument.
```{r, fig.dim=c(8, 6)}
aevolcano(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention", p_value_cutoff=0.05)
```

The default colours for the bubbles are blue and red, with blue indicating greater risk in the control arm and red indicating greater risk in the intervention arm. We can change the colour of bubbles through the `bubble_colours` argument, specifying the colour indicating greater risk in the control arm first, followed by the colour indicating greater risk in the intervention arm.
```{r, fig.dim=c(8, 6)}
aevolcano(df2, body_system_class="aebodsys", control="Placebo", intervention="Intervention", bubble_colours = c("#B2E061", "#FFB55A"))
```

## Saving tables and plots
### Docx
We can save tables as docx by specifying the filepath in the `save_docx_path` argument.
```{r, fig.dim=c(8, 6)}
aetable(df2, body_system_class="aebodsys", control="Placebo", intervention_levels=c("Intervention"),
        treatment_effect_estimate=TRUE, variables = c("variable1", "variable2"), save_docx_path="aetable.docx")
```

### Image
We can save tables and plots as images by specifying the filepath in the `save_image_path` argument.
```{r, fig.dim=c(8, 6)}
aebar(df3, arm_levels=c("Intervention 1","Intervention 2", "Placebo"), arm_names=c("Drug A","Drug B", "No Drug"),
      adverse_event="ae_pt", facets=TRUE, save_image_path="aeimage.png")
```

